---
date: "`r Sys.Date()`"
---

 

::: {.alert .alert-info}
## **Objetive** {.unnumbered .unlisted}

-   Demonstrate how to obtain acoustic data from online repositories
:::

 

```{r, echo=FALSE}

source("~/Dropbox/R_package_testing/suwo/R/map_locations.R")
```


<font size="4">

## Getting recordings from online repositories

The suwo package offers several functions to query and download recordings (and other media) of animal vocalizations in online repositories. For instance, the following code return recordings metadata from Xeno-Canto:

Example on how to get recording's metadata (no downloading):

```{r, message=FALSE,  warning=FALSE}

library(suwo)

phae_his <- query_xenocanto(term = 'Phaethornis hispidus')

```

Keep only song vocalizations:

```{r}

song_phae_his <- phae_his[grep("song", ignore.case = TRUE, phae_his$vocalization.type), ]

```

Map locations using `map_locations()`:

```{r, fig.width=10, echo=TRUE}

map_locations(song_phae_his)

```

Once you feel fine with the subset of data you can go ahead and download it as follows:

```{r, eval = FALSE, fig.width=10}

download_media(metadata = song_phae_his, 
         path = "DIRECTORY WHERE YOU WANT TO SAVE THE FILES")

```

------------------------------------------------------------------------

<font size="4">Session information</font>

```{r session info, echo=F}

sessionInfo()

```
